Party tricks for PostgreSQL \\ perf, ftrace and bpftrace

### 2022-10-24 pgconfeu2022/presentation.tex
Adjustments

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome5}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usetikzlibrary{spy, decorations, decorations.pathmorphing, decorations.pathreplacing, backgrounds, decorations.text}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}
\usetikzlibrary{arrows.meta}

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{itemize}[label={}]
            \item {Static tracing}
            \item {Tracepoints}
            \item {USDT}
        \end{itemize}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{itemize}[label={}]
            \item {Dynamic tracing}
            \item {kprobe/kretprobe}
            \item {uprobe/uretprobe}
        \end{itemize}
    \end{center}
\end{frame}

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ bpftrace -p 31715 -f json -e
    'usdt:bin/postgres:postgresql:checkpoint__start
    {printf("Checkpoint, Flags %d\n", arg0);}'

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large, escapeinside=::]{c}
*dovacuum =
	force_vacuum ||
	(vactuples > vacthresh) ||
	(vac_ins_base_thresh >= 0 &&
		:\colorbox{red!20}{instuples > vacinsthresh)};:
        \end{minted}
    \end{center}
\end{frame}
\note{
	Filter doesn't work with comparisons, requires post-processing,
	complicated. For such cases it's good to have a stable usdt.
}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{itemize}[label={}]
            \item<+-> {Context switch overhead}
            \item<+-> {Debugging symbols}
            \item<+-> {Incorporate into infrastructure?}
            \item<+-> {Privileges?}
        \end{itemize}
    \end{center}
\end{frame}

@@

hitcount: 1
{ lat_us: 314,
  query_string: select b from test2;}
        \end{minted}
    \end{center}
\end{frame}

@@

\begin{frame}
    \frametitle{}
    \begin{center}
        \includegraphics[width=0.9\textwidth,center]{more.png}
    \end{center}
\end{frame}

@@

# ========
# captured on    : Wed Oct 12 17:07:32 2022
# header version : 1
# data offset    : 504
# data size      : 1392968
# feat offset    : 1393472
# arch : x86_64
# cpudesc : Intel(R) Core(TM) i7-10610U CPU @ 1.80GHz
# ...
        \end{minted}
    \end{center}
\end{frame}
\note{ENOMEM, only for those functions that are markes safe to fail.}

@@

\begin{frame}
    \frametitle{}
    \begin{center}
        \includegraphics[width=0.7\textwidth,center]{elephant.jpg}
    \end{center}
\end{frame}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome5}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usetikzlibrary{spy, decorations, decorations.pathmorphing, decorations.pathreplacing, backgrounds, decorations.text}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{copyrightbox}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}
\usetikzlibrary{arrows.meta}

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        Static tracing: Tracepoints, USDT
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        Dynamic tracing: k[ret]probe, u[ret]probe
    \end{center}
\end{frame}

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ bpftrace -p <pid> -f json -e
    'usdt:bin/postgres:postgresql:checkpoint__start
    {printf("Checkpoint, Flags %d\n", arg0);}'

@@

%\begin{frame}[fragile]{}
    %\frametitle{}
    %\begin{center}
        %\begin{minted}[fontsize=\large, escapeinside=::]{c}
%*dovacuum =
	%force_vacuum ||
	%(vactuples > vacthresh) ||
	%(vac_ins_base_thresh >= 0 &&
		%:\colorbox{red!20}{instuples > vacinsthresh)};:
        %\end{minted}
    %\end{center}
%\end{frame}
%\note{
	%Filter doesn't work with comparisons, requires post-processing,
	%complicated. For such cases it's good to have a stable usdt.
%}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{itemize}[label={}]
            \item<+-> {Context switch overhead}
            \item<+-> {Debugging symbols}
            \item<+-> {Post-processing}
            \item<+-> {Incorporate into infrastructure?}
            \item<+-> {Privileges?}
        \end{itemize}
    \end{center}
\end{frame}

@@

hitcount: 1
{ lat_us: 314,
  query_string: select b from test2;}
        \end{minted}
    \end{center}
\end{frame}
\note{
    visual hists if possible
}

@@

\renewcommand\UrlFont{\color{gray!50}\rmfamily\itshape}
\begin{frame}
    \frametitle{}
    \begin{center}
        \copyrightbox[b]
            {\includegraphics[width=0.9\textwidth,center]{more.png}}
            {Belzebubs: \url{belzebubsofficial.tumblr.com}}
    \end{center}
\end{frame}
\note{
    Copyright, link
}

@@

# ========
# captured on    : Wed Oct 12 17:07:32 2022
# header version : 1
# data offset    : 504
# data size      : 1392968
# feat offset    : 1393472
# arch : x86_64
# ...
        \end{minted}
    \end{center}
\end{frame}
\note{ENOMEM, only for those functions that are markes safe to fail.}

@@

\begin{frame}
    \frametitle{}
    \begin{center}
        \copyrightbox[b]
            {\includegraphics[width=0.7\textwidth,center]{elephant.jpg}}
            {Dmitrii Dolgov: Teufelsberg, Berlin}

    \end{center}
\end{frame}
\note{
    Copyright, name place
}
